---
title: "NHANES Challenge"
format: html
editor: visual
---

### Load Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(tidyverse)
library(haven)
library(here)
library(ggplot2)
library(janitor)

```

### Read-in Datasets for this Analysis:

-   DR1TOT_J.xpt as 'diet_behavior'
-   2017-2018_Hg-Blood.XPT as 'blood_hg'
-   2017-2018_Hg-Urine.XPT as 'urine_hg'
-   2017-2018_Diabetes.XPT as 'diabetes'

```{r}
diet_behavior <- read_xpt(here("data/nhanes_data/DR1TOT_J.XPT"))
blood_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Blood.XPT"))
urine_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Urine.XPT"))
diabetes <- read_xpt(here("data/nhanes_data/2017-2018_Diabetes.XPT"))
demographics <- read_xpt(here("data/nhanes_data/2017-2018_Demographics.XPT"))
```

### Subset Read-in Datasets

Subset 'diet_behavior' as 'diet'

```{r}
diet <- select(diet_behavior, SEQN, DRD360, DRD370B, DRD370BQ, DRD370Q, DRD370QQ)
```

Subset 'diabetes' as 'tiid'

```{r}
tiid <- select(diabetes, SEQN, DIQ010)
```

Subset 'blood_hg' as 'bhg'

```{r}
bhg <- select(blood_hg, SEQN, LBXIHG, LBDIHGSI, LBXBGE, LBXBGM)
```

Subset "urine_hg' as 'uhg'

```{r}
uhg <- select(urine_hg, SEQN, URXUHG)
```

### Merge Subsets Into A Working Dataframe as 'df'

```{r}
df <- list(diet, tiid, bhg, uhg)

df <- df %>% reduce(full_join, by = 'SEQN')
```

1.  Filter Dataframe df for the following:

```{r}
# Assuming your dataframe is named `nhanes_data`
df <- df %>%
  # Filter out rows where DIQ010 or DRD360 are NA
  filter(!is.na(DIQ010), !is.na(DRD370B)) %>%
  # Keep only rows where DIQ010 and DRD360 are 1 or 2
  filter(DIQ010 %in% c(1, 2), DRD370B %in% c(1, 2)) %>%
  # Recode 1 to "Yes" and 2 to "No" for DIQ010 and DRD360
  mutate(
    DIQ010 = ifelse(DIQ010 == 1, "Has Diabetes", "No Diabetes"),
    DRD370B = ifelse(DRD370B == 1, "Consumes Ahi", "No Ahi")
  )
  view(df)
```

2.  Do some exploratory data analysis
3.  Determine the prior probability that someone has diabetes (DIQ010)
    -   P(B)
4.  Determine the prior probability that someone eats tuna (DRD370B)
    -   P(A)
5.  Determine the likelihood of eating tuna and having diabetes
    -   L(B\|A)
6.  Determine posterior probability via bayes theorem
    -   P(B\|A)
7.  What other questions can we ask of this data?

[**Bayes Theorem**]{.underline}

B \<- Has Diabetes

A \<- Consumes Ahi

P(B)

The probability that a person has diabetes

```{r}
df %>%
  tabyl(DIQ010)
```

```{r}
probability_has_diabetes <- 0.13
```

L(B\|A) = P(A\|B)

```{r}
df%>%
  tabyl(DRD370B, DIQ010)%>%
  adorn_percentages("col")
```

```{r}
probabilty_consumes_ahi_given_diabetes <- 0.36 
```

P(A)

```{r}
df%>%
  tabyl(DRD370B)
```

```{r}
probability_consumes_ahi <- 0.36
```

P(B\|A)

```{r}
(probability_has_diabetes * probabilty_consumes_ahi_given_diabetes)/probability_consumes_ahi
```

```{r}
demographics <- demographics %>%
  select(SEQN, RIDRETH1, RIAGENDR, RIDAGEYR)
```

```{r}
nhanes_data <- left_join(df, demographics, by = "SEQN")
```

```{r}
nhanes_data <- nhanes_data %>%
  mutate(age_bracket = ifelse(RIDAGEYR > 40, "old", "young"))
```

```{r}
ggplot(nhanes_data, aes(x = RIDAGEYR, fill = DIQ010)) + geom_density(alpha = 0.5)
```

P(B\|A)

B \<-

```         
```

```{r}
over_40 <- nhanes_data %>%
  filter(RIDAGEYR > 40)

```

```{r}
over_40 %>%
  tabyl(DIQ010)
```

```{r}
probability_has_diabetes_over_40 <- 0.23
```
